---
title: "RepData_PeerAssessment2"
author: "Janus"
date: "Thursday, September 17, 2015"
output: html_document
---

#####An Analysis Report on the Impact of Severe Weather Events on Public Health and Economy in the United States

#####Data processing 

```{r}
setwd("C:/Users/ITSUPPORT/Desktop/R/A2")
library(R.utils)
library(ggplot2)
library(plyr)
require(gridExtra)
#####Read the data.

```{r}
stormData <- read.csv("repdata-data-StormData.csv", sep = ",")
dim(stormData)
```
#####The time period of the events begins in the year 1950 and ends in November 2011.
```{r}
if (dim(stormData)[2] == 37) {
    stormData$year <- as.numeric(format(as.Date(stormData$BGN_DATE, format = "%m/%d/%Y %H:%M:%S"), "%Y"))
}
hist(stormData$year, breaks = 30)
```
#####According to the histogram, the number of events significantly increases around 1995. By utilizing the subset of the data from 1990 to 2011.
```{r}
storm <- stormData[stormData$year >= 1995, ]
dim(storm)
```
After subsetting, there are 681500 rows and 38 columns.

####Impact on Public Health

#####We notices at the number of fatalities and injuries that are caused by several weather events.

```{r}
sortHelper <- function(fieldName, top = 15, dataset = stormData) {
    index <- which(colnames(dataset) == fieldName)
    field <- aggregate(dataset[, index], by = list(dataset$EVTYPE), FUN = "sum")
    names(field) <- c("EVTYPE", fieldName)
    field <- arrange(field, field[, 2], decreasing = T)
    field <- head(field, n = top)
    field <- within(field, EVTYPE <- factor(x = EVTYPE, levels = field$EVTYPE))
    return(field)
}

```
```{r}
fatalities <- sortHelper("FATALITIES", dataset = storm)

```
```{r}
injuries <- sortHelper("INJURIES", dataset = storm)
```

####Impact on Economy

#####We will convert the property damage and crop damage data into comparable numerical forms according to the meaning of units described in the code book. 
```{r}
convertHelper <- function(dataset = storm, fieldName, newFieldName) {
    totalLen <- dim(dataset)[2]
    index <- which(colnames(dataset) == fieldName)
    dataset[, index] <- as.character(dataset[, index])
    logic <- !is.na(toupper(dataset[, index]))
    dataset[logic & toupper(dataset[, index]) == "B", index] <- "9"
    dataset[logic & toupper(dataset[, index]) == "M", index] <- "6"
    dataset[logic & toupper(dataset[, index]) == "K", index] <- "3"
    dataset[logic & toupper(dataset[, index]) == "H", index] <- "2"
    dataset[logic & toupper(dataset[, index]) == "", index] <- "0"
    dataset[, index] <- as.numeric(dataset[, index])
    dataset[is.na(dataset[, index]), index] <- 0
    dataset <- cbind(dataset, dataset[, index - 1] * 10^dataset[, index])
    names(dataset)[totalLen + 1] <- newFieldName
    return(dataset)
}

storm <- convertHelper(storm, "PROPDMGEXP", "propertyDamage")
```
```{r}
storm <- convertHelper(storm, "CROPDMGEXP", "cropDamage")
```
```{r}
names(storm)
```
```{r}
options(scipen=999)
property <- sortHelper("propertyDamage", dataset = storm)
crop <- sortHelper("cropDamage", dataset = storm)
```

####Results

#####Looking at the impact on public health, there are two sorted lists of severe weather events shown below by the number of people badly affected.

```{r}
fatalities
```
```{r}
injuries
```
#####According to the graphs shows the total fatalities and injuries resulting from serveral weather events.

```{r}
fatalitiesPlot <- qplot(EVTYPE, data = fatalities, weight = FATALITIES, geom = "bar", binwidth = 1) + 
    scale_y_continuous("Number of Fatalities") + 
    theme(axis.text.x = element_text(angle = 45, 
    hjust = 1)) + xlab("Severe Weather Type") + 
    ggtitle("Total Fatalities by Severe Weather\n Events in the U.S.\n from 1995 - 2011")
injuriesPlot <- qplot(EVTYPE, data = injuries, weight = INJURIES, geom = "bar", binwidth = 1) + 
    scale_y_continuous("Number of Injuries") + 
    theme(axis.text.x = element_text(angle = 45, 
    hjust = 1)) + xlab("Severe Weather Type") + 
    ggtitle("Total Injuries by Severe Weather\n Events in the U.S.\n from 1995 - 2011")
grid.arrange(fatalitiesPlot, injuriesPlot, ncol = 2)
```
#####According to the histograms, flood and hurricane/typhoon result in the most property damage, while drought and flood result in the most crop damage in the United States from 1995 to 2011.

####Conclusion

#####From the analysis, we conclude that excessive heat and tornado are most harmful with respect to population health, while flood, drought, and hurricane/typhoon have the greatest economic consequences.
